---
title: "Weighted Race Lottery"
author: "Garret Christensen"
date: "September 17, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## 2020 Race Lottery Background

Thanks to all who entered. The selection was conducted randomly using a computer program written by Garret Christensen. The script is available on GitHub [here](http://www.github.com/garretchristensen/RaceLottery).

```{r seed}
set.seed(1492)
```
The seed for the pseudo-random number generator was set using 4 rolls of a 10-sided die, held in public at Law's Whiskey Bar in Salida, CO on Nover 15, 2019.

## 2020 Race Lottery Winners
The data file of entrants in the lottery, including the characteristics used in the lottery (previous finishes, previous starts, previous applications) can also be found in the GitHub repository [here](http://www.github.com/garretchristensen/RaceLottery). 
```{r results, echo=FALSE}
df<-read.csv("./fakedata.csv")
head(df)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
