---
title: "Weighted Race Lottery"
author: "Garret Christensen"
date: "October 1, 2019"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(tibble)
library(readxl)

hl2019<-read_excel("./HiLo Entrants by year.xlsx")
hl2018<-read_excel("./HiLo Entrants by year.xlsx", sheet = "2018")
hl2017<-read_excel("./HiLo Entrants by year.xlsx", sheet = "2017")

ccc<-read_excel("./2019 CCC Lottery data scraped for High Lonesome lottery testing.xlsx")

temp<-read.csv("./fakedata.csv", stringsAsFactors = FALSE) #LOAD THE DATA
df<-as_data_frame(temp)
```
# Testing Different Lottery Setups
We have High Lonesome and Cascade Crest data to work with. We'll make an educated guess about how many HL applicants we'll have, and run a handful of versions of the lottery based on that. We'll also use the CCC data and see what mix of entrants we'd get with their data.

```{r HLdata, echo=FALSE}
men2017=nrow(men<-hl2017[which(hl2017$gender=="M"),])
men2018=nrow(men<-hl2018[which(hl2018$gender=="M"),])
men2019=nrow(men<-hl2019[which(hl2019$gender=="M"),])
women2017=nrow(women<-hl2017[which(hl2017$gender=="F"),])
women2018=nrow(women<-hl2018[which(hl2018$gender=="F"),])
women2019=nrow(women<-hl2019[which(hl2019$gender=="F"),])

```
## Predicting the Future
Is impossible, but let's guess how many HL applicants there will be for 2020:

Male entrants: `r men2017` in 2017, `r men2018` in 2018, and `r men2019` in 2019.

Female entrants: `r women2017` in 2017, `r women2018` in 2018, and `r women2019` in 2019.

Male applicant growth rate was `r round(((men2019/men2018)-1)*100, digits=1)`% in 2019 and `r ((men2018/men2017)-1)*100`% in 2018, so let's assume that stays basically constant at `r round(((men2019/men2018)-1)*100, digits=1)`% and assume `r round(men2019/men2018*men2019, digits=0)` male entrants for 2020. Female applicant growth rate was was `r round(((women2019/women2018)-1)*100, digits=1)`% in 2019 and `r round(((women2018/women2017)-1)*100, digits=1)`% in 2018. That's a big decline in the growth rate, but maybe the new lottery will get women excited to enter? Let's guess the growth rate will go back up a little, but not all the way (1/3 the first year's growth, 2/3 the second year's, or `r round(((1/3*women2018/women2017 + 2/3*women2019/women2018)-1)*100, digits=1)`%). That's `r round(women2019*(1/3*women2018/women2017 + 2/3*women2019/women2018), digits=0)` female applicants.

Since this is more applicants than we have previously observed, I'll randomly re-sample the existing 2019 applicants to make up the difference.

## Lottery Design Choices
There's no perfect design, but we have a few values we'd like to try and implement.

* We want equal numbers of men and women (statiscally this is easy)
* We'd like to get a mix of new and veteran runners, without guaranteeing entry for either
* Previous unsuccessful applications should be nearly as valuable as running
* We value volunteering and trail work
* We'd like new entrants to have a decent chance to run within a couple-few years

There's an infinite number of ways to design this, but here are some basic choices:

* Linear--just add up points
* Exponential--weight certain items by 2^N (or really, anthing greater than 1)
* Caps--limit the number of tickets from a certain column (e.g. if you've run it 5+ times, you can let somebody else get a turn)



# Running the Lottery
The remainder of this document implements a single version of the lottery, formatting the output as you might want to for the actual event.
## 2020 Race Lottery Background

Thanks to all `r nrow(df)` who entered. The selection was conducted randomly using a computer program written by Garret Christensen. The script is available on GitHub [here](http://www.github.com/garretchristensen/RaceLottery).

```{r seed, echo=FALSE, results=TRUE}
set.seed(1494) #SET THE SEED WITH DICE!
n<-125 #SET THE TOTAL NUMBER TO PICK

#NUMBER OF MEN AND WOMEN APPLICANTS
n_men_app=nrow(men<-df[which(df$female==0),])
n_women_app=nrow(women<-df[which(df$female==1),])

#WHO HAS MORE APPLICANTS AND WOULD GET THE ODD SLOT, IF APPLICABLE?
 if (n_men_app > n_women_app) {
   largest<-"men"
 } else {
     largest<-"women"
 }

#FIRST, DO WE EVEN HAVE TO BOTHER RUNNING A LOTTERY?
if (n_men_app+n_women_app<n) {
  print ("THERE IS NO REASON TO RUN A LOTTERY! Everyone gets in!")
  n_women_pick<-n_women_app
  n_men_pick<-n_men_app
  knitr::knit_exit()
} else {

#IS THE ENTRY CAP EVEN?
#IF SO, WHICH SEX HAS MORE APPLICANTS?
#AND DO WOMEN HAVE ENOUGH APPLICANTS?
if((n %% 2) == 0) {
  #print("Entry cap is even.")
    #CAP IS EVEN
    #DO WOMEN HAVE ENOUGH APPLICANTS?
    if (n_women_app>=n/2) {
        #print("Women have enough applicants to fill their race.")
        #YES THEY DO
        n_women_pick<-n/2
        n_men_pick<-n/2
    }   else {
         #print("Women don't have enough applicants to fill their race.")
        #NO THEY DON'T, GIVE EXTRA SLOTS TO MEN
        n_women_pick<-n_women_app
        n_men_pick=n-n_women_pick
    }
} 

#IS THE ENTRY CAP ODD?
#IF SO, WHICH SEX HAS MORE APPLICANTS?
#AND DO WOMEN HAVE ENOUGH APPLICANTS? 
#NOTE WE DON'T CHECK IF MEN HAVE ENOUGH--IT'S ASSUMED.
if ((n %% 2) == 1) {
  #print("Entry cap is odd.")
  if (n_women_app>n/2) {
  #print("Women have enough applicants to fill their race.")
    
  if (n_men_app>n_women_app) {  
    #print("Men have more applicants and get the rounding slot.")
    n_men_pick<-round((n/2+.00001), digits = 0)
    n_women_pick<-round((n/2)-1, digits = 0)
  } else {
    #print("Women have at least equal applicants and get the rounding slot.")
    n_women_pick<-round((n/2+.00001), digits = 0)
    n_men_pick<-round((n/2)-1, digits = 0)
  } #END WOMEN HAVE ENOUGH
  } else {
    #print("Women don't have enough applicants to fill their race.")
    n_women_pick<-n_women_app
    n_men_pick=n-n_women_pick
  }  
}

} # END OVERALL NEED TO RUN LOTTERY
```
The seed for the pseudo-random number generator was set using 4 rolls of a 10-sided die, held in public at Law's Whiskey Bar in Salida, CO on Nover 15, 2019.

The number of slots available through the lottery is: `r n`. Half the slots go to men and half to women. (If there are an odd number of slots, whichever sex has more entrants gets the extra slot, ties going to women. In our case, that's `r largest`. `r n_women_app` women applied and `r n_men_app` men applied, so we'll be picking `r n_women_pick` women and `r n_men_pick` men.)

## 2020 Race Lottery Entrants
The data file of all `r nrow(df)` entrants in the lottery, including the characteristics used in the lottery (previous finishes, previous starts, previous applications) can also be found in the GitHub repository [here](http://www.github.com/garretchristensen/RaceLottery). 

Just to give you an idea, the dataset looks like this:
```{r entrants, echo=FALSE}
df<-df[,c(1,2,4,5,7,8,9,3)] #rearrange. 6 was temporary/made-up
head(df)
```

And summary characteristics of entrants are as follows:

`r mean(df$female)*100`% Female.

Average Age: `r round(mean(df$age), digits=2)`

Average Finishes: `r round(mean(df$finishes), digits=2)`

Average Starts: `r round(mean(df$starts), digits=2)`

Average Applications: `r round(mean(df$applications), digits=2)`


## 2020 Race Lottery Entrants
Entrants were given one ticket for each year of the following: starting, finishing, and applying for the race. The winners are listed below, and their names have been saved to CSV files, ./women_winners.csv and ./men_winners.csv. 

The mens and womens lotteries were conducted separately, with equal number of men and women chosen, except in the (unlikely, we hope!) event of fewer women entering than slots available. In that case, we reassigned extra slots to men.

The female lottery winners are:
```{r winners, echo=FALSE}
#stupid example formula for a persons tickets
#dplyr function sample_n will work with weights, normalize automatically
#syntax:sample_n(tbl, size, replace = FALSE, weight = NULL, .env = NULL, ...)
df$tickets<-df$finishes+df$starts+df$applications+1

#make the mens and womens pools (separate tibbles)
#newdata <- mydata[ which(mydata$gender=='F' 
women<-df[which(df$female==1),]
men<-df[which(df$female==0),]

#Run the separate lotteries
women_winners<-sample_n(women, n_women_pick, replace = FALSE, weight=women$tickets)
men_winners<-sample_n(men, n_men_pick, replace = FALSE, weight=men$tickets)

#Print the winners' names
write.csv(women_winners$fullname)
```
The male lottery winners are:
```{r men, echo=FALSE}
write.csv(men_winners$fullname)
```

```

